Time Series Modeling of Varying Complexity Captures Subdiffusive Solute Dynamics and Predicts Long Timescale Behavior in Nanoscale Pores.

### 2020-01-01 Ben_Manuscripts/stochastic_transport/Draft.tex
Better explanations of beta

  Anti-correlation between solute hops has a significant effect on the scaling
  of solute flux with pore length. In Figure~\ref{fig:beta_hurst}, we show that 
  $\beta$, as derived from the sFBMcut model, is inversely related to the Hurst
  parameter while $\beta$, as derived from the MSDDM, shows no discernible
  trend.

  \begin{figure}
  \centering
  \begin{subfigure}{0.43\textwidth}
  \includegraphics[width=\textwidth]{beta_hurst.pdf}
  \caption{}\label{fig:beta_hurst}
  \end{subfigure}
  \begin{subfigure}{0.52\textwidth}
  \includegraphics[width=\textwidth]{beta_parameters.pdf}
  \caption{}\label{fig:beta_parameters}
  \end{subfigure}
  \end{figure}

@@

  % Partially in methods?
  The power law decay of the flux with pore length implies the following relationship
  for selectivity via substitution of Equation~\ref{eqn:flux_decay} into 
  Equation~\ref{eqn:selectivity}:
  \begin{equation}
  S_{ij}(L) = \bigg(\frac{c_i}{c_j}\bigg)L^{(\beta_i - \beta_j)}
  \label{eqn:selectivity_ratio}
  \end{equation}

  %MRS5: figure seem to show that the selectivity plateaus more or less, so thin coatings enough. 
  In Figure~\ref{fig:selectivity}, we plot Equation~\ref{eqn:selectivity_ratio} for pore
  lengths ranging from those studied in Figure~\ref{fig:flux_curves} to macroscopic-length
  pores. Greater differences in $\beta$ values leads to selectivities that are stronger 
  functions of pore length. For $(c_i / c_j)=1$, LLC membranes will be more selective
  towards passage of solutes with lower $\beta$ values versus those with high $\beta$ values.
  When $\beta$ values are equal, as is the case for uncorrelated motion, selectivity only
  depends on the pre-factors, $c$. 
  %MRS5? can you remind people of the physical source (in terms of dwell times, hop lengths) of the differences in beta and c? Want to try to stablish the connection between particle dynamic behavior and selectivity.

  \begin{figure}
  \centering
  \includegraphics[width=0.5\textwidth]{selectivity.pdf}
  \caption{The selectivity of 2 species changes monotonically with pore length. The
  strength of dependence on pore length depends on the difference between $\beta$ values.}
  \label{fig:selectivity}
  \end{figure}

@@

  \section*{Acknowledgments}
  %MRS: add people who contributed non-author level ideas. 
  %MRS3: add specific info for PRF and GAANN
  %BJC3: I don't know the PRF info. I wrote GAANN as I did on the last paper which is what Rob Davis told me to do
  %MRS4: I willō track down the PRF grant number (if typical)
  This work was supported in part by the ACS Petroleum Research Fund grant \#59814-ND7
  and the Graduate Assistance in Areas of National Need (GAANN)
  fellowship which is funded by the U.S. Department of Education.
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.

>>>>

  %The scaling of solute flux with pore length is a complicated function of hop and 
  %dwell parameters, however, anti-correlation between solute hops appears to
  %be the most significant influencing factor. 
  The scaling of solute flux with pore length is primarily influenced by anti-correlation
  between solute hops. In Figure~\ref{fig:beta_hurst}, we show that $\beta$, as derived 
  from the sFBMcut model, is inversely related to the Hurst parameter. This makes intuitive
  sense since higher degrees of anti-correlation should slow the rate at which solutes
  cross the membrane pore. $\beta$, as derived from the MSDDM, appears approximately 
  constant but this is consistent with nearly invariant transitional Hurst parameters
  (see Figure~\ref{fig:H_v_state}). However, Figure~\ref{fig:beta_hurst} does raise
  the question as to why $\beta$ is so much lower than the sFBMcut model.

  To understand the behavior of the MSDDM, we studied the $\beta$ parameter derived 
  from variations on each model, visualized in Figure~\ref{fig:beta_parameters}. First, 
  we removed anti-correlation between hops in the sFBMcut trajectory (set $H$=0.5). 
  The $\beta$ parameter drops to values similar to the MSDDM. When we remove anti-correlation
  \textit{and} dwell times between hops (set dwell times equal to 1), effectively simulating 
  Brownian motion, we observe similar $\beta$ parameters. These two variations imply that,
  in the AD models, the $\beta$ parameter is only a function of $H$ and not dependent on
  dwell times ($\alpha_d$) or hop lengths ($\sigma$). The MSDDM is fundamentally similar to 
  the AD model in that one makes correlated draws from a hop distribution. Therefore,
  we should expect the $\beta$ value of the MSDDM to be larger than that of Brownian 
  motion.
  %Although the dwell times are
  %insignificant compared to the AD model (compare Figures~\ref{fig:ad_eyetest} and
  %\ref{fig:msddm_eyetest} for example), we have just shown

  The value of $\beta$ for the MSDDM is under-estimated due to assumptions of the model
  itself as well as inaccuracies resulting from of simulation of very long FLM trajectories.
  Turning first to the model itself, we have designed it to ensure that the magnitude of the
  hops in the series of transitions between states are anti-correlated from start to finish. 
  This assumes that the transition correlation structure is unaffected by the
  sub-trajectories between each state transition. Each time a state transition occurs, one
  must initialize a new time series sub-trajectory with its own correlation structure. 
  Even with a low Hurst parameter, as exhibited by most trapped states, the time spent 
  in each state is relatively short making the accuracy of the correlation structure 
  questionable. Additionally, $\sigma$ of transitional hops is similar in magnitude to
  $\sigma$ in trapped states. Since we add the transitional hop lengths to each end of
  trapped state sub-trajectories, the transitional hop lengths are shifted with respect 
  to one another largely decorrelating them. We tested this reasoning by modifying the 
  MSDDM to completely immobilize particles except when they transition between states. 
  Suprisingly, $\beta$ is still close to the Brownian value. Further experimentation 
  reveals that this is actually a consequence of the FLM simulation procedure. Simulating
  FLM requires Riemann-sum approximations of the stochastic integrals defining the process.
  In short, it is computationally infeasible to use enough terms to accurately incorporate
  long timescale correlations into our long MSDDM realizations. Thus at long timescales, 
  we lose correlation between transitional jumps. We confirmed this hypothesis by using 
  fractional Brownian motion, for which we have an exact simulation method, in place of 
  FLM. When we use FBM, $\beta$ increases well above the Brownian value. $\beta$ increases
  even further if we immobilize the trapped states. Thus the low value of $\beta$ of the
  MSDDM is a consequence of model assumptions and inexact simulation of FLM.

  \begin{figure}
  \centering
  \begin{subfigure}{0.43\textwidth}
  \includegraphics[width=\textwidth]{beta_hurst.pdf}
  \caption{}\label{fig:beta_hurst}
  \end{subfigure}
  \begin{subfigure}{0.52\textwidth}
  \includegraphics[width=\textwidth]{beta_parameters.pdf}
  \caption{}\label{fig:beta_parameters}
  \end{subfigure}
  \caption{(a) The $\beta$ values of the sFBMcut model appear to be inversely proportional
  to the Hurst parameters. The $\beta$ values of the MSDDM are much lower with much less
  variation than the sFBMcut model. (b) The $\beta$ parameters of the MSDDM are low because
  sub-trajectories between state transitions decorrelate transitional hops and because
  our FLM simulation procedure does not accurately correlate hops after long time lags. 
  The high $\beta$ parameters of the sFBMcut model are a consequence of anti-correlation
  between hops. Removing hop correlations causes $\beta$ to drop down close to MSDDM
  values (sFBMcut ($H$=0.5)). Removing dwell times in addition to hop correlation
  yields a similar value of $\beta$ (Brownian). Immobilizing particles while in a trapped state
  also yields a similar value of $\beta$ (MSDDM (immobile states)). Replacing FLM with
  FBM in the MSDDM raises $\beta$ well above the Brownian value (MSDDM (FBM)). Replacing FLM with FBM and
  immobilizing particles in trapped states further raises the value of $\beta$.}\label{fig:beta}
  \end{figure}

  The values of $c$ are likely a complex function of $\sigma$, $alpha_d$, $H$ and $\alpha_d$
  where applicable. The values of $c$ scale the flux curves and are directly correlated to 
  flux magnitude. We showed earlier that $\beta$ is not a function of $\sigma$ or $\alpha_d$,
  therefore dwell times and hop lengths must influence $c$. In fact, any parameter which acts
  to modify the rate at which solutes move through the membrane pores will affect $c$.
  Therefore it is sensible to hypothesize that a higher degree of anti-correlation (lower $H$)
  should decrease $c$ and a higher value of $\alpha_h$ should also increase $c$.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{c_parameter_comparison.pdf}
  \caption{}\label{fig:c_parameters}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{c_parameter_comparison.pdf}
  \caption{}\label{fig:c_influence}
  \end{subfigure}
  \caption{(a) The c paramters are directly related to your mom (b) }\label{fig:c}
  \end{figure}

@@

  The power law decay of the flux with pore length implies the following relationship
  for selectivity via substitution of Equation~\ref{eqn:flux_decay} into 
  Equation~\ref{eqn:selectivity}:
  \begin{equation}
  S_{ij}(L) = \bigg(\frac{c_i}{c_j}\bigg)L^{(\beta_i - \beta_j)}
  \label{eqn:selectivity_ratio}
  \end{equation}

  %MRS5: figure seem to show that the selectivity plateaus more or less, so thin coatings enough. 
  In Figure~\ref{fig:selectivity}, we plot Equation~\ref{eqn:selectivity_ratio} for pore
  lengths ranging from those studied in Figure~\ref{fig:flux_curves} to macroscopic-length
  pores. Greater differences in $\beta$ values leads to selectivities that are stronger 
  functions of pore length. For $(c_i / c_j)=1$, LLC membranes will be more selective
  towards passage of solutes with lower $\beta$ values versus those with high $\beta$ values.
  When $\beta$ values are equal, as is the case for uncorrelated motion, selectivity only
  depends on the pre-factors, $c$. In most cases, the length dependence of selectivity 
  plateaus quickly which implies that a coating of LLC membrane with a thickness on the 
  order of those already made experimentally should be sufficiently optimal. 
  %MRS5? can you remind people of the physical source (in terms of dwell times, hop lengths) of the differences in beta and c? Want to try to stablish the connection between particle dynamic behavior and selectivity.

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{selectivity.pdf}
  \caption{sFBMcut}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{selectivity_msddm.pdf}
  \caption{MSDDM}
  \end{subfigure}
  \caption{The selectivity of 2 species changes monotonically with pore length. The
  strength of dependence on pore length depends on the difference between $\beta$ values.
  The selectivities predicted by the sFBMcut AD model (a) are much larger in magnitude than
  those predicted by the MSDDM (b) because predictions of the AD model result in a much
  larger variation in flux (compare Figures~\ref{fig:flux_curves_ad} and~\ref{fig:flux_curves_msddm}).}
  \label{fig:selectivity}
  \end{figure}

  The selectivities predicted by the sFBMcut model are both higher in magnitude and
  ordered differently than the those predicted by the MSDDM. The magnitudes of the
  sFBMcut predictions are higher because the predicted solute fluxes vary more widely
  than those predicted by the MSDDM (compare Figures~\ref{fig:flux_curves_ad} 
  and~\ref{fig:flux_curves_msddm}). The ordering is different mostly because the 
  ordering of the flux predictions is different.

  We cannot say with absolute certainty which selectivity curves are more 
  trustworthy. The sFBMcut model uses exact simulation techniques and yields 
  qualitatively accurate solute trajectories, but it may overestimate long 
  timescale correlation present in the real system. For this reason, we might
  expect the $\beta$ values of the true flux curves to be closer to the 
  Brownian value as exhibited by the MSDDM. But the MSDDM yields qualitatively
  inaccurate solute trajectories and we do not have a good understanding of
  how to precisely control the point at which solute hops become decorrelated.

  However, we may gain the most by using the two models in tandem. In both cases,  
  the data suggests that this particular LLC membrane might be most useful for 
  selectively separating ethylene glycol from acetic acid.

@@

  \section*{Acknowledgments}
  %MRS: add people who contributed non-author level ideas. 
  %MRS3: add specific info for PRF and GAANN
  %BJC3: I wrote GAANN as I did on the last paper which is what Rob Davis told me to do
  This work was supported in part by the ACS Petroleum Research Fund grant \#59814-ND7
  and the Graduate Assistance in Areas of National Need (GAANN)
  fellowship which is funded by the U.S. Department of Education.
  Molecular simulations were performed using the Extreme Science and
  Engineering Discovery Environment (XSEDE), which is supported by National
  Science Foundation grant number ACI-1548562. Specifically, it used the Bridges
  system, which is supported by NSF award number ACI-1445606, at the Pittsburgh
  Supercomputing Center (PSC). This work also utilized the RMACC Summit supercomputer,
  which is supported by the National Science Foundation (awards ACI-1532235 and
  ACI-1532236), the University of Colorado Boulder, and Colorado State
  University. The Summit supercomputer is a joint effort of the University of
  Colorado Boulder and Colorado State University.

  We would also like to thank Richard Noble for helping us make the connection 
  between the empirical mean first passage time distribution and the analytical
  equation to which we fit the distributions (Equation~\ref{eqn:passage_times}).
